\thetitle

### 2020-11-21 writing/introduction.tex
almost done with results

The paper's second contribution is a set of results from applying the evaluation
method to three distinct blame assignment strategies and approximately
72000 cases (section~\ref{sec:results}): Vitousek et
al.'s strategy of tracking typed/untyped boundaries in Reticulated;
Tobin-Hochstadt et al.'s use of higher-order contract system~\cite{ff-icfp-2002,
mf-toplas-2009}; and the approach of industrial systems that forego blame
and instead rely on error messages from the
underlying programming language. A direct interpretation of the results
(section~\ref{sec:discussion}) is that
they validate the conjecture behind the work of theoreticians. A good blame assignment
strategy helps the search for impedance mismatches between the specified types
and the behavior of untyped components. Even better, the evaluation method 
identifies the wrapper-based blame tracking of Typed Racket as the most
useful academic system with that of Reticulated Python just a short
distance behind.

>>>>

The paper's second contribution is a set of results from applying the evaluation
method to three distinct blame assignment strategies and approximately
72000 cases (section~\ref{sec:results}): Vitousek et
al.'s strategy of tracking typed/untyped boundaries in Reticulated;
Tobin-Hochstadt et al.'s use of higher-order contract system~\cite{ff-icfp-2002,
mf-toplas-2009}; and the approach of industrial systems that forego blame
and instead rely on error messages from the
underlying programming language. A direct interpretation of the results
(section~\ref{sec:discussion}) is that
they validate the conjecture behind the work of theoreticians. A good blame assignment
strategy helps the search for impedance mismatches between the specified types
and the behavior of untyped components. Even better, the evaluation method 
identifies the wrapper-based blame tracking of Typed Racket as the most
useful academic system with that of Reticulated Python only marginally  behind.

### 2020-11-21 writing/results.tex
almost done with results

\begin{figure*}
  \centering
  \includegraphics[width=0.7\textwidth]{./plots/usefulness-table}
  \caption{Usefulness: Each cell depicts the proportion of a benchmark's mutants for which a mode is always useful (green), sometimes useful (yellow), and never useful (red).}
  \label{fig:usefulness-table}
\end{figure*}

@@

\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{./plots/avo-matrix}
  \caption{Usefulness comparisons: Each cell depicts the percentage of a
  benchmark's mutants for which the row mode is more useful than the column mode.}
  \label{fig:avo-matrix}
\end{figure}

We present our results according to the three dimensions for assessing the utility of blame described in section~\ref{sec:rational}: how often blame is always, sometimes, and never useful (figure~\ref{fig:usefulness-table}); programmer effort (figure~\ref{fig:effort-table}); and how often one mode adds value over another mode (figure~\ref{fig:avo-matrix}).
In this section we analyze each of these dimensions in turn, and in the next we discuss the analysis's high level conclusions.

Figure~\ref{fig:usefulness-table} breaks down how often blame is useful (as defined in section~\ref{sec:rational}) for every benchmark and every mode of the rational programmer.
Each mode corresponds to a row of the table, while the columns correspond to our benchmarks.
Each cell plots the proportion of a benchmark's mutants for which a mode is always useful (green), sometimes useful (yellow), and never useful (red).
Comparing the overall pattern of one row to the others provides a broad understanding of how the usefulness of that mode compares with the others.
For instance, the large proportions of red across Erasure indicates that for many mutants it is never useful.

Figure~\ref{fig:effort-table} shows the distribution of trail lengths for every benchmark and every mode of the rational programmer.
The modes correspond to rows and columns to benchmarks, like figure~\ref{fig:usefulness-table}.
In this case, however, the cells plot trail length distributions starting from length zero in the leftmost bar and going up by one for each bar to the right.
Each bar is broken into two sub-bars, which are color-coded to indicate the proportion of trails within that bar which are successful (green) and not (red).
Again, comparing the rows of the table provides a high level view of how the lengths of trails for each mode tend to differ.
For example, the stark difference between the Null mode's distribution and that of every mode illustrates how using blame information (as opposed to randomly selecting components to type) dramatically shortens the rational programmer's search.
The Null distributions look approximately normal \emdash reflecting that the Null mode always selects a random next component to type \emdash but shifted left, reflecting that our scenarios start with a randomly-selected subset of components already typed.
On the other hand, all of the other modes have distributions that are significantly more concentrated around trails of length one and two.

Finally, figure~\ref{fig:avo-matrix} shows a head-to-head comparison of each mode against every other mode.
Each cell of the table compares one mode to another using section~\ref{sec:rational}'s definition of a mode \emph{adding value over} another mode.
In particular, a cell in row $R$ and column $C$ plots the percentage of mutants for which $R$ adds value over $C$.
For example, the second cell in the row of Natural shows that Natural adds value over Natural exceptions for 18.9\% of all the mutants we consider.
Reading across a given row $R$ provides an overview of how $R$ compares to every other mode;
the higher the bars, the more mutants for which $R$ is strictly more useful than the other modes.
Reading down a given row $C$ provides an overview of how every other mode compares to $C$;
the higher the bars, the more mutants for which $C$ is strictly less useful than the other modes.

>>>>

\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{./plots/avo-matrix}
  \caption{Usefulness comparisons: Each cell depicts the estimated percentage of
  interesting debugging scenarios for which the row mode is more useful
  than the column mode.
  The upper bound of error margin is 0.02\%}
  \label{fig:avo-matrix}
\end{figure}

We run the experimental process from section~\ref{sec:rational} on the
sampled debugging scenarios from section~\ref{sec:sample} on a --- specs
of fix. Each debugging scenario had a timeout limit of 4 minutes and an
out of memory limit of 6GB. The complete experiment required
approximately 270 hours of compute hours on our machine.

Figure~\ref{fig:avo-matrix} summarizes our results and provides answers to
the four questions from section~\ref{subsec:experiment}.  Specifically, it
depicts a head-to-head comparison of the usefulness of each mode against
every other mode.  The cell in row $R$ and column $C$ plots the percentage
of interesting debugging scenarios for which $R$ is more useful $C$ ---
this percentage has been calculated as an estimated proportion from the
corresponding proportion of the debugging scenarios in our sample and the
upper bound.  In effect, reading across a given row $R$ provides an
overview of how $R$ compares to every other mode; the higher the bars, the
more interesting debugging scenarios for which $R$ is strictly more useful
than the other modes.  Reading down a given row $C$ provides an overview
of how every other mode compares to $C$; the higher the bars, the more
interesting debugging scenarios for which $C$ is strictly less useful than
the other modes. For example, the second cell in the row of Natural blame
shows that Natural is more useful than Natural exceptions for 12\% of all
the interesting debugging scenarios we generate from our benchmarks and
mutators.

Figure~\ref{subsec:experiment} shows that the answers for questions $Q_1$
to $Q_3$ are positive. In all three semantics, blame modes outperform
their corresponding exception mode (by 9.44\% to 12.7\%). The natural
exceptions mode is never more useful than natural blame and  transient
exceptions are more useful than transient first and transient last blame
in 2.19\% and 2.8\% of the scenarios respectively.

For the answers to $Q_*$, we compare the percentages in antisymmetric
positions in the matrix. Blame for all three semantics is significantly
more useful than Erasure exceptions (by 33\% to 35\%). Natural blame is
more useful than both versions of transient blame by a small percentage;
in 4.05\% of scenarios natural blame is more useful while in 1.67\% of the
scenarios transient blame is more useful. The transient first and
transient last blame are practically indistinguishable. Finally, there is
no clear winner between natural exceptions and transient exceptions
despite the theoretically advantageous additional checks of the natural
semantics.

@@

Figure~\ref{fig:effort-table} shows the distribution of programmer effort
for the sampled interesting debugging --- in contrast to the usefulness
comparisons these results do not generalize to the full population.

### 2020-11-21 writing/sample.tex
almost done with results

The mutators from section~\ref{sec:mutate} generate 96,515,744 debugging
scenarios; far too many to even identify the interesting ones among them.
Furthermore this population is heterogeneous; different scenarios come
from different mutants and different mutants are the result of a
different mutator and benchmark pair. Hence, to make the experiment
computationally feasible and statistically precise, we perform stratified
uniform random sampling with three levels of stratification.  The first
two strata group mutants first by benchmark and then by mutator. 
We randomly select from the mutants of
each benchmark ensuring that, per
benchmark, the sample reflects the diversity of mutants with respect to
the mutators that generated them.  Specifically,  we sample 80 mutants
with interesting scenarios per benchmark,  evenly-distributed across all
of the mutators that contribute mutants for the benchmark.  The third
stratum of our sampling is the lattice of a mutant and we draw 96 random
interesting scenarios from each lattice. For lattices that have less than
96 interesting scenarios, we select them all.

In summary, our sampling process takes advantage of ancillary information
to deliver a  sufficiently large representative set of scenarios.
Concretely, we test each
mode of the rational programmer on 72,192 randomly selected interesting
debugging scenarios spread across both benchmarks and mutators.
Furthermore stratified random sampling allows us to derive precise conclusions
about the proportion of the whole population that satisfies a given trait 
from the proportion of the sample 
that satisfies the trait despite the population's heterogeneity, exactly what we need to project comparisons about the
usefulness of various modes of the
rational programmer from the sample to the set of all interesting debugging
scenarios.  Finally, we can also calculate separate estimates per stratum
and provide an alternative
viewpoint of our results. For instance, because of the size of the sample
of scenarios from the lattice of a mutant, we can estimate whether natural
blame is more useful than natural exceptions for all interesting scenarios
of each mutant with 95\% confidence and 5\% error margin.

>>>>

The mutators from section~\ref{sec:mutate} generate 96,515,744 debugging
scenarios; far too many to even identify the interesting ones among them.
Furthermore this population is heterogeneous; different scenarios come
from different mutants and different mutants are the result of a
different mutator and benchmark pair. Hence, to make the experiment
computationally feasible and statistically precise, we perform stratified
uniform random sampling with three levels of stratification.  The first
two levels of strata group mutants first by benchmark and then by mutator. 
We randomly select from the mutants of
each benchmark ensuring that, per
benchmark, the sample reflects the diversity of mutants with respect to
the mutators that generated them.  Specifically,  we sample 80 mutants
with interesting scenarios per benchmark,  evenly-distributed across all
of the mutators that contribute mutants for the benchmark.  The third
level of strata for our sampling is the lattice of a single mutant and we draw 96 random
interesting scenarios from each lattice with replacement.

In summary, our sampling process takes advantage of ancillary information
to deliver a  sufficiently large representative set of scenarios.
Concretely, we test each
mode of the rational programmer on 72,192 randomly selected interesting
debugging scenarios spread across both benchmarks and mutators.
Furthermore stratified random sampling allows us to derive precise conclusions
about the proportion of the whole population that satisfies a given trait 
from the proportion of the sample 
that satisfies the trait despite the population's heterogeneity, exactly what we need to project comparisons about the
usefulness of various modes of the
rational programmer from the sample to the set of all interesting debugging
scenarios.
